---
title: "BioSys"
author: "Florian Mayer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r libs, message=FALSE}
library(biosysR)
```

# Setup
The BioSys API is only accessible with basicauth using a valid Biosys username 
and password.

All `biosysR` functions calling the BioSys API expect optional parameters `un` 
and `pw`, which default to environment variables `BIOSYS_UN` and `BIOSYS_UN`,
respectively. Calling `biosysR` functions with invalid or empty, credentials, or
defaulting to non existing `BIOSYS_UN` and `BIOSYS_PW` will fail with an 
informative error message prompting for correct authentication.

There are three ways to supply these authentication credentials to `biosysR` functions.

## Permanent authentication
To set-and-forget BioSys authentication, add to your `~/.Rprofile`:

```{r setenv, eval=FALSE}
Sys.setenv(BIOSYS_UN = "USERNAME")
Sys.setenv(BIOSYS_PW = "PASSWORD")
```

Every new R session will already contain these variables.

## Session authentication
To authenticate one session, export `BIOSYS_UN/PW` as environment variables:

```{r setenv_session, eval=FALSE}
Sys.setenv(BIOSYS_UN = "USERNAME")
Sys.setenv(BIOSYS_PW = "PASSWORD")
```

Restarting the R session will clear these variables.

## Per request authentication
Supply the variables to each `biosysR` function:

* `biosys_projects`
* `biosys_datasets`
* `biosys_records`

```{r setenv_request, eval=FALSE}
projects <- biosys_projects(un="USERNAME", pw="PASSWORD")
```

Doing so will hand `un` and `pw` to `biosys_get`, which builds the authentication
headers and uses them in the request to BioSys.

# Data flow

* Benthic images are analysed and annotated in the software EcoPAAS
* EcoPAAS outputs data as Excel
* BioSys imports Excel using a config created by Paul's uploader (tm)
* [Biosys GUI](https://biosys.dbca.wa.gov.au/)
* [BioSys API in Browser](https://biosys.dbca.wa.gov.au/sso/api) 
using single sign on (needs active browser window)
* [BioSys API in scripts](https://biosys.dbca.wa.gov.au/api) using basicauth 
(username and password)
* [BioSys API documentation](https://biosys-admin.dbca.wa.gov.au/api/explorer/)

# Accessing data from BioSys
## Helper functions

* Data is retrieved from the BioSys API through an HTTP GET with basicauth.
* The JSON returned from the BioSys API is parsed into a `tibble`.
* All heavy lifting is factored out into helper functions.

## BioSys projects
```{r projects}
projects <- biosys_projects()
dplyr::glimpse(projects)
DT::datatable(projects)
```

## BioSys datasets
```{r datasets}
datasets <- biosys_datasets(project_id = 6)
dplyr::glimpse(datasets)
listviewer::jsonedit(datasets$data_package)
DT::datatable(datasets)
```

## BioSys records
```{r records}
records <- biosys_records(project_id = 6)
DT::datatable(head(records, n = 100))
```

# Example data
The example data in this project were produced by saving one project's
project metadata, dataset metadata, and a subset of records to the `data/` folder.

```{r generate_example_data, eval=FALSE}
projects <- biosys_projects()[6,]
datasets <- biosys_datasets(project_id = 6)[1,]
records <- biosys_records(project_id = 6) %>% head(n = 100)
devtools::use_data(projects, datasets, records)
```